# Experimental design {#sec-design}

```{r}
#| label: setup
#| warning: false
suppressMessages(library(here)) # for pathing to/from project directory
suppressMessages(library(tidyverse)) # for dplyr, ggplot, etc.
suppressMessages(library(janitor)) # for cleaning data
suppressMessages(library(glue)) # for cleaning data

# source scripts
clean <- readRDS(here("data", "clean.rds"))
full <- readRDS(here("data", "full.rds"))

source(here("R", "helpers.R"))
source(here("R", "plots.R"))
```

## Key takeaways

- **Most studies focus on syntax**, not surprisingly. In many cases, the line between morphology and syntax or syntax and semantics is a bit blurry, but this is not really a main focus.
- **Explicit reporting of design features**, such as exact numbers of conditions, **is not very consistent**. 
- **Studies vary widely in the numbers of conditions** they include, and many studies, especially earlier ones, report overly complex designs. 

## Linguistic levels in focus

```{r}
#| label: fig-phenom
#| fig-cap: "Proportions of studies focusing of levels of linguistic structure"
#| fig-width: 8
#| fig-height: 4
make_split_perc_gg_barplot(clean, "linguistic_phenomenon",
                           title = "by level of linguistic focus")
```


## Condition specification

:::{.panel-tabset}

### Reporting of condition numbers

Was the **total number of conditions explicitly reported**?

The exact number of experimental conditions is not consistently reported, although the design can often be reconstructed from the description. Studies often report individual factors and/or design structure, e.g. "We used a 2x3 design crossing factor A with factor B", but then leave the reader to compute the final tally of experimental conditions (6 in this example). We coded any case where the *exact number is not strictly stated* as "Not Reported", thus the data here are more conservative than one might expect. 

```{r}
#| label: fig-des-perc-rep
#| fig-cap: "Proportions of studies reporting the exact numbers of experimental conditions"
#| fig-height: 2.5
#| fig-width: 8
clean |> 
  mutate(N_conditions_reported = str_replace_all(N_conditions_reported, "No", "Not Reported")) |> 
  make_perc_gg_barplot("N_conditions_reported", title = "reporting exact condition numbers")
```

In some instances certain contrasts are not considered part of the experimental conditions, thus under-representing the complexity of the design.


### Conditions

How many **experimental conditions** are there?

There are `r length(is.na(clean$N_conds))` studies where the total number of conditions is not clear. Those with larger numbers (>10) should probably be re-examined. There are even a few studies (N = `r length(which(clean$N_conds >= 50))`) with very large (>50) numbers of conditions, which should certain be checked and confirmed.

Again, many of these numbers are deducted / inferred from incomplete descriptions of designs. We've included some examples of unclear descriptions in the [Examples](examples.qmd) page. 

```{r}
#| label: fig-cond-boxplot
#| fig-cap: "There is a surprisingly wide range of experimental conditions included in AJ studies"
#| lightbox: false
make_boxplot(clean, "N_conds", xlab = "Experimental Conditions", xrange = c(0, 50))
```

### Design structure

Did the study use a **within-subjects** or **between-subjects** design?

Most studies are not interested in between-groups patterns.

```{r}
#| label: fig-design-type
#| fig-cap: "Most studies use within-subjects designs"
#| fig-height: 3.5
#| fig-width: 8
make_perc_gg_barplot(clean, "design_type", title = "using within- or between-subjects design")
```

:::

## Additional design characteristics

:::{.panel-tabset}

### Randomization

Were items presented in a **fixed or (semi-)random order** across participants?

- *Random*: fully random for each participant
- *Semi-random* ('psuedo-random): order is largely random, but with some prescribed conditions on order (e.g. no adjacent items of the same condition)
- *Fixed*: order is the same across (sets of) participants. Fixed orders are probably more common, as paper questionnaires do not allow for participant-specific re-ordering, but this is not always reported explicitly.

Surprisingly this is not clearly reported in the texts. It's possible the "Not Reported" numbers are inflated due to inconsistent description (randomization is often mentioned cursorily in unexpected places in texts). 

```{r}
#| label: fig-randomization
#| fig-cap: "A surprising number of studies fail to report how items were ordered"
#| fig-width: 8
#| fig-height: 3.5
make_perc_gg_barplot(clean, "randomization", title = "reporting of presentation order")
```

### Presentation

How were the **experimental trials organized** for presentation?

- *Blocks*: items are arranged into and present as predetermined sets 
- *Semi-random* ('psuedo-random): order is largely random, but with some prescribed conditions on order (e.g. no adjacent items of the same condition)

Here "intermixed" involves items of different conditions being intermixed without blocking. Blocking typically also involves intermixing within blocks. Often this is not described clearly at all.

```{r}
#| label: fig-presentation
#| fig-cap: "Most studies intermix items without blocking"
#| fig-width: 8
#| fig-height: 3
make_perc_gg_barplot(clean, "exp_blocking", title = "reporting of item organization")
```


### Other features

- Some studies (N = `r nrow(filter(clean, N_pre.test != "0"))`) include **pre-test trials**, typically as practice trials for the task. This is common with Magnitude Estimation studies. Very few (N = `r nrow(filter(clean, N_post.test != "0"))`) include **post-test trials**.
- Many studies balanced items across (predicted) conditions, but again this is not always reported clearly.
- A small percentage of studies involve the same items in multiple tasks.

```{r}
#| label: tbl-design-details
#| tbl-cap: "Reporting rates for other design details"
clean |>
  mutate(
    n_pre = ifelse(`N_pre.test` > 0, "Yes", "No"),
    n_post = ifelse(`N_post.test` > 0, "Yes", "No"),
  ) |> 
  select(balanced_items, n_pre, n_post, multiple_tasks) |>
  pivot_longer(everything(), names_to = "Detail", values_to = "value") |>
  mutate(
    Detail = case_when(
      Detail == "balanced_items" ~ "Stimuli balanced across conditions?",
      Detail == "n_pre" ~ "Included pre-test trials?",
      Detail == "n_post" ~ "Included post-test trials?",
      Detail == "multiple_tasks" ~ "Same items used across multiple judgment tasks?"
    ),
    value = ifelse(value %in% c("Not Reported", "", "No"), "No", "Yes")
  ) |>
  count(Detail, value) |>
  pivot_wider(names_from = value, values_from = n, values_fill = 0) |>
  mutate(
    perc_reported = round(100 * Yes / (Yes + No), 1)
    ) |>
  rename("No / Not Reported" = "No", `% Yes` = perc_reported) |>
  knitr::kable()
```


:::

