---
title: "Overview"
---

```{r}
#| label: setup
suppressMessages(library(here)) # for pathing to/from project directory
suppressMessages(library(tidyverse)) # for dplyr, ggplot, etc.
suppressMessages(library(janitor)) # for cleaning data

# source scripts
raw <- readRDS(here("data", "raw.rds"))
clean <- readRDS(here("data", "clean.rds"))
full <- readRDS(here("data", "full.rds"))

source(here("R", "helpers.R"))
source(here("R", "plots.R"))

# Global variables
total_articles <- nrow(full)
n_coded <- length(unique(raw$article_index))
perc_coded <- round(100*n_coded/total_articles, 1)
journals <- unique(full$journal)
n_languages <- clean |> 
  mutate(
    language = str_replace_all(language, "(/| and )", "; "),
    language = str_remove_all(language, " -.*")
  ) |> 
  pull(language) |> 
  unique() |> 
  length()

filtered <- raw |> 
  filter(!str_detect(coder_comments, "[Ee]xclude|No judgments collected")) 

n_excluded <- length(unique(raw$article_index)) - length(unique(filtered$article_index))
```


This is just a (not so) quick overview of where we stand with coding, and some observations about what we've found so far. 

Coding has been a slog to say the least...

## Where we are with annotations 

### By Journal

So far we've annotated **`r paste0(n_coded,"/",total_articles, " (", perc_coded, "%)")` articles**, containing a total of `r nrow(raw)` individual acceptability judgment task (AJT) studies. Some of these (N = `r n_excluded`) were exculded because they didn't actually include formal AJT experiments.

::: {#fig-art-by-journal}
```{r}
#| lightbox: false
make_stacked_perc_barplot(full, "journal")
```

We've still got just a few journals to finish!
:::

### By year

```{r}
#| label: fig-art-by-year
#| fig-cap: "Numbers of AJs in these journals have been increasing since the 2010s."
#| lightbox: false
make_stacked_perc_barplot(full, "year", sort = F, orientation = "v")
```


## Annotation contribution

Most of the coding was done by JG!

```{r}
#| label: tbl-coder-summary
#| tbl-cap: "Articles coded by coder"
raw |>
  distinct(article_index, coder_id) |> 
  mutate(coder_id = toupper(coder_id)) |>
  count(coder_id, name = "Articles coded", sort = T) |>
  rename(Name = "coder_id") |> 
  knitr::kable()
```


## Languages investigated

We've identified `r n_languages` language varieties studied in our review thus far. We'll need to double check the "Not Reported" cases. I don't think we should be applying a strict version of the ["Bender Rule"](https://thegradient.pub/the-benderrule-on-naming-the-languages-we-study-and-why-it-matters/) here, but we could do so at a later point. This will mostly apply to the English studies.

```{r}
#| label: fig-languages
#| fig-cap: "English is by far the most commonly studied language in the sample"
#| fig-width: 8
#| fig-height: 10
clean |> 
  mutate(
    language = str_replace_all(language, "(/| and )", "; "),
    language = str_remove_all(language, " -.*")
  ) |> 
  make_split_gg_barplot("language", base_size = 12)
```
















