# Participants {#sec-participants}

```{r}
#| label: setup
suppressMessages(library(here)) # for pathing to/from project directory
suppressMessages(library(tidyverse)) # for dplyr, ggplot, etc.
suppressMessages(library(janitor)) # for cleaning data

# source scripts
clean <- readRDS(here("data", "clean.rds"))
full <- readRDS(here("data", "full.rds"))

source(here("R", "helpers.R"))
source(here("R", "plots.R"))
```


## Key takeaways

- We find that **most studies involve fewer than 60-70 participants** (mean = `r mean(as.numeric(clean$N_parts), na.rm = T)`, median = `r median(as.numeric(clean$N_parts), na.rm = T)`), but there are some with huge numbers, for example @sprouse_gender_2021 included `r max(clean$N_parts, na.rm = T)` participants. 
- The **population that participants are assumed to be sampled from is rarely explicitly stated**.
- University affiliation (students and/or staff) has traditionally been the most common way to get participants, but online platforms have become more common in recent studies. 
- **Compensation is not reported very often**, although online platforms often pay participants directly, so this may be assumed by authors. But it should be explicitly reported (and often is).
- Not all studies report demographics, but usually at least L1 is specified.
- **Power analysis is almost never reported / conducted**, see [here](@sec-power).

## Numbers

How many **total participants** were recruited?

We've had several studies (N = `r sum(is.na(as.numeric(clean$N_participants_recruited)))`) for which the number of participants was not fully clear.

```{r}
#| label: fig-part-num
#| fig-cap: "Numbers of participants vary, but usually are less than 100"
#| lightbox: false
make_boxplot(clean, "N_parts", xlab = "Participants Recruited", xrange = c(0, 410))
```


## Participant characteristics

::: {.panel-tabset}

### Population

What was the **population** that participants were sampled from? 

The population from which participants are sampled is not always clear. University students are common sources of participants, but details are often minimal. Populations of online platforms like AMT are not well known, although many studies do report controls used for narrowing recruitment, such as limiting IP addresses to certain countries. Populations from these platforms are coded as "Mixed".

```{r}
#| label: fig-part-pop
#| fig-cap: "Proportions of populations reported across studies"
#| fig-width: 8
#| fig-height: 4
make_perc_gg_barplot(clean, "part_profile", base_size = 14,
                     title = "by participant population")
```

### Demographics

Which participant **demographics** were reported?

Not all studies report demographics, but usually at least being a native speaker ("Language Background") is specified. Any mention of average age or sex proportions counted as reporting for our purposes.  

```{r}
#| label: fig-part-demo
#| fig-cap: "Proportions of demographics reported across studies"
#| fig-width: 8
#| fig-height: 4
make_split_perc_gg_barplot(clean, "demographics_reported", base_size = 14,
                           title = "by demographic information reported")
```

:::

## Recruitment and compensation

:::{.panel-tabset}

### Recruitment

How were participants **recruited**?

University affiliation (students and/or staff) is the most common way to get participants, but online platforms have become much more common as we annotate more recent studies. 

Nonetheless, many studies continue to fail to report how participants were recruited. Even those involving students do not consistently describe *how* involvement from students was solicited (e.g. via fliers, social media, classrooms, word-of-mouth). 

```{r}
#| label: fig-part-rec
#| fig-cap: "Proportions of recruitment methods reported across studies"
#| fig-width: 8
#| fig-height: 4.5
make_perc_gg_barplot(clean, "part_recruitment", title = "by recruitment method")
```

### Filtering

Did study report how many participants were included in the **final analysis after filtering**?

Most studies do not explicitly report whether they removed data from any participants after initial recruitment.

```{r}
#| label: fig-filter
#| fig-cap: "Proportions of studies reporting whether participants were filtered/excluded from final analysis"
#| fig-height: 2
#| fig-width: 8
clean |> 
  mutate(
    N_parts_filtered_bin = ifelse(is.na(N_parts_filtered_bin), FALSE, N_parts_filtered_bin),
    N_parts_filtered_bin = ifelse(N_parts_filtered_bin, "Reported", "Not Reported")
    ) |> 
  make_perc_gg_barplot("N_parts_filtered_bin", title = "reporting participant filtering")
```


### Compensation

How were participants **compensated**?

Compensation is not often reported, particularly when off-line recruitment methods are used. Online platforms such as AMT and Prolific pay participants directly, so this may be assumed by authors. Many do report payments, but this is not always explicitly reported. Studies occasionally do explicitly report *not* compensating participants.

```{r}
#| label: fig-part-comp
#| fig-cap: "Proportions of compensation methods reported across studies"
#| fig-width: 8
#| fig-height: 4
make_perc_gg_barplot(clean, "compensation", title = "by participant compensation")
```



:::
