# Stimuli {#sec-stimuli}

```{r}
#| label: setup
suppressMessages(library(here)) # for pathing to/from project directory
suppressMessages(library(tidyverse)) # for dplyr, ggplot, etc.
suppressMessages(library(janitor)) # for cleaning data
suppressMessages(library(glue)) # for cleaning data

# source scripts
clean <- readRDS(here("data", "clean.rds"))
full <- readRDS(here("data", "full.rds"))

source(here("R", "helpers.R"))
source(here("R", "plots.R"))
```


## Key takeaways

- **Explicit reporting of item numbers is not very consistent**. The number of **items per condition** is rarely explicitly mentioned, nor is the number of participants who see a given item. Reconstruction suggests that for many studies the number of judgments per item varies across items. 
- The choice of **the number of items used is almost never explained or justified**, e.g. via power analysis.
- Detailed **information about fillers is limited**. Studies often report the number of fillers, and some general descriptions, but as with critical items almost no studies report *why they chose that number*. Also, when item sets are provided, filler items are almost never included.

## Item Reporting

:::{.panel-tabset}

### Explicit reporting

Was the **total number of items explicitly reported**?

A surprising number of articles do not accurately report item numbers, where we use the term "item" to refer to a single unique stimulus token that a participant is intended to respond to (see [Glossary](@ssec-item-definition)). Similar to the number of experimental conditions, exact numbers can often be reconstructed / deduced from information in the article, with varying degrees of effort and certainty. So again, the data here are conservative counts.

The primary issue is that use of the term "item" is not consistent across studies. Many authors use it to denote sets of sentences that are balanced across experimental conditions. For instance, @frazier_anticipating_2018 describe their experiment like so:

> "Sixteen items like the one illustrated in (6) were constructed. **Each item had four versions**, defined by the factorial combination of +NR versus −NR and the use of a verb that permitted Neg Raising versus one that did not [...]" (emphasis added)
 
In our coding this *should* be reported as 16 x 4 = 64 unique test sentences, hence 64 "items". This final number is not explicitly reported, hence we code this as "No" (total item number NOT reported).

```{r}
#| label: fig-items-reported
#| fig-cap: "Proportion of studies reporting the exact numbers of stimulus items"
#| fig-width: 8
#| fig-height: 2
clean |> 
  count(N_items_reported) |> 
  mutate(
    N_items_reported = ifelse(N_items_reported == "Yes", "Reported", "Not Reported")
  ) |> 
  make_perc_gg_barplot("N_items_reported", title = "reporting exact numbers of items")
```




to refer to a single unique stimulus that a participant is intended to respond to

### Total items

What was the total number of **critical items** created for the study *not* including fillers or distractors?

There are `r length(is.na(clean$N_items))` studies where the total number of items is not clear. Additionally we have `r length(which(clean$N_items >= 200))` studies with 200 or more items. most of these very large studies are replication studies or comparisons with frequency data, e.g. @white_frequency_2020 who included 50000 items!

```{r}
#| label: fig-item-boxplot
#| fig-cap: "Distribution of critical item counts across studies"
#| warning: false
#| lightbox: false
clean |> 
  make_boxplot("N_items", xlab = "Critical items", xrange = c(0, 420))
```

On the lower end, there are some studies that report semi-formal surveys of small sets of items, or even just one item, as in @boas_coercion_2011a. These are often reported in footnotes with minimal details.

### Items per condition

How many **critical items are there per experimental condition**?

The number of items per condition (min = 1, median = `r median(clean$N_items_per_conds, na.rm = T)`, max = `r max(clean$N_items_per_conds, na.rm = T)`) is an important aspect of study design, as it directly relates to the power of the study. Frustratingly, this is not always reported clearly.

We find `r nrow(filter(clean, str_detect(N_items_per_condition, "^\\d", negate=T)))` studies did not provide clear information about item numbers per condition. 

```{r}
#| label: fig-item-per-cond-boxplot
#| fig-cap: "Distribution of critical item counts per condition across studies"
#| warning: false
#| lightbox: false
clean |>
  mutate(
    N_items_per_condition = as.numeric(N_items_per_condition),
    N_items_per_condition = ifelse(is.na(N_items_per_condition), 0, N_items_per_condition)
  ) |>
  filter(N_items_per_condition>0, N_items_per_condition<100) |> 
  # make_histogram("N_items_per_condition")
  make_boxplot("N_items_per_condition", xlab = "Items per condition")
```

### Fillers 

What was the number of **filler or distractor** items?

We find that `r nrow(filter(clean, N_fillers %in% c("0", "Not Reported")))` do not mention including filler items.

```{r}
#| label: fig-fillers
#| fig-cap: "Distribution of critical item counts per condition across studies"
#| warning: false
clean |>
  mutate(
    N_fillers = as.numeric(N_fillers),
    N_fillers = ifelse(is.na(N_fillers), 0, N_fillers)
  ) |>
  filter(N_fillers>0) |> 
  # make_histogram("N_items_per_condition")
  make_boxplot("N_fillers", xlab = "Number of filler items")
```


:::

## Additional stimuli characteristics

:::{.panel-tabset}

### Construction

Were the **items constructed by the researchers** or based on authentic materials?

This is not always explicitly mentioned, but it usually can be inferred from the text.

```{r}
#| label: fig-item-construciton
#| fig-cap: "Most items are constructed by researchers, but this is also not always reported clearly"
#| fig-width: 8
#| fig-height: 3.5
make_perc_gg_barplot(clean, "item_construction")
```

### Controls

Were the **items balanced/controlled for internal features** that were *not* the focus of investigation, e.g. frequency, length, etc.?

```{r}
#| label: fig-item-controls
#| fig-cap: "Non-focal factors are rarely controlled for in item design"
#| fig-width: 8
#| fig-height: 4
make_split_perc_gg_barplot(clean, "item_balance")
```

### Context

Were items presented with **additional context**?

The "Other" category here may involve any kind of phenomenon or language specific factors related to lexical items, pragmatics, etc. This code probably could be more fine-grained. 

```{r}
#| label: fig-item-context
#| fig-cap: "Most items are constructed by researchers, but this is also not always reported clearly"
#| fig-width: 8
#| fig-height: 3
make_perc_gg_barplot(clean, "contextualization")
```

### Piloting

Were **items pilot-tested or normed** in any way?

A handful of studied explicitly report piloting or norming items, but no studies explicitly report NOT doing so.

```{r}
#| label: fig-items-pilot
#| fig-cap: "Studies rarely report running any kind of **norming or pilot testing** of critical items"
#| fig-height: 2.5
#| fig-width: 8
make_perc_gg_barplot(clean, "item_plausibility")
```

:::



















