# Judgment task {#sec-task}

```{r}
#| label: setup
suppressMessages(library(here)) # for pathing to/from project directory
suppressMessages(library(tidyverse)) # for dplyr, ggplot, etc.
suppressMessages(library(janitor)) # for cleaning data
suppressMessages(library(glue)) # for cleaning data

# source scripts
clean <- readRDS(here("data", "clean.rds"))
full <- readRDS(here("data", "full.rds"))

source(here("R", "helpers.R"))
source(here("R", "plots.R"))
```

## Key takeaways

- **Likert-type scales dominate**: about 7 in 10 studies use an ordinal (Likert) response scale, most often with 7 points or 5 points. Binary (yes/no, grammatical/ungrammatical) judgments are the next most common format.
- **The construct being rated is often unclear**. "Acceptability" is the most common term used to describe the judgment by the authors, but "Not Reported" is the *second* most common category — meaning readers often have to guess whether participants were rating grammaticality, naturalness, or something else entirely.
- **Task administration details are rarely reported**: timing (self-paced vs. fixed), whether sessions were supervised, whether there were practice/pilot trials, breaks, or comprehension checks are almost never mentioned.
- Delivery format has shifted toward computer-based administration, but a substantial share of studies still don't report much detail on how the task was delivered.

## Response format

::: {.panel-tabset}

### Scale type

What was the **scale** of participants' responses?

```{r}
#| label: fig-response-scale
#| fig-cap: "Ordinal (Likert) scales are by far the most common response format."
#| fig-width: 8
#| fig-height: 4
make_perc_gg_barplot(clean, "response_scale", base_size = 12)
```

### Scale levels

How many **levels of acceptability** were used (2, 7, 100, …)?

For studies using an ordinal scale of 3 or more levels, the number of response levels is fairly concentrated around 7 and 5 points.

```{r}
#| label: fig-response-levels
#| fig-cap: "Distribution of response scale lengths (Likert scales of 3 or more only)"
#| fig-width: 8
#| fig-height: 4
plot_rating_levels_gg(clean)
```

### Rating terminology

What **terminology** did the authors use to refer to their judgment task with participants (e.g. "rate how acceptable / natural /… each sentence is")? 

An important note is that **most articles do not provide the exact instructions given to participants**, and only offer paraphrases. In at least one case [@francis_grammatical_2010a], the authors' paraphrase in the text was inconsistent with the actual instructions seen by participants. 

```{r}
#| label: fig-rating-terminology
#| fig-cap: "Terminology used to describe the judgment being elicited"
#| fig-width: 8
#| fig-height: 4.5
make_split_perc_gg_barplot(clean, "rating_terminology", base_size = 12)
```

The label used matters. "Acceptability," "grammaticality," and "naturalness" are not interchangeable constructs, yet a suprising number of studies (N = `r nrow(filter(clean, rating_terminology == "Not Reported"))`) don't specify which one they're measuring [we need to check this].


:::

## Presentation and delivery

::: {.panel-tabset}

### Modality

What **modality** was the judgment task delivered in?

Again, this is often not reported explicitly, but it can be reasonably inferred from the text.

```{r}
#| label: fig-presentation-modality
#| fig-cap: "Written presentation dominates; auditory AJTs remain relatively rare"
#| fig-width: 8
#| fig-height: 3.5
make_perc_gg_barplot(clean, "presentation_modality")
```

### Delivery format

How were participant **responses** recorded?

Any studies using software tools and online delivery were automatically coded as "Computer-based". In many other cases it's not clear how participants entered responses.

```{r}
#| label: fig-delivery-format
#| fig-cap: "Computer-based delivery is now the most common format, though paper surveys persist"
#| fig-width: 8
#| fig-height: 3.5
make_perc_gg_barplot(clean, "delivery_format")
```

:::

## Task administration

Beyond the response format itself, details of *how* the task was administered — timing, supervision, practice, and attention checks — matter for interpreting results, but are inconsistently reported.

::: {.panel-tabset}

### Duration

Was the **task duration** recorded?


Many studies do report average performance time, and most experimental software and platforms record time automatically. But yet again---say it with me now---this is not often explicitly reported in the text.

```{r}
#| label: fig-task-duration
#| fig-cap: "Did the study mention how long participants took (usually as an average)?"
#| fig-width: 8
#| fig-height: 3
make_perc_gg_barplot(clean, "task_duration")
```


### Timing

Were items presented in a **fixed or variable amount of time**, or was the presentation **self-paced** (participants proceded at their own pace)? 

Some studies may use speeded judgment tasks. Most allow participants to proceed at their own pace but this is rarely explicitly stated.

```{r}
#| label: fig-task-timing
#| fig-cap: "Whether the task was self-paced or had a fixed presentation time is rarely reported"
#| fig-width: 8
#| fig-height: 3
make_perc_gg_barplot(clean, "task_timing")
```

### Supervision

Was the study conducted under **supervision of the researcher** (e.g. in a lab or classroom with researcher present)?

Any study administered in a lab or classroom was coded as "Yes". Studies using online crowd-sourcing platforms (AMT, Prolific) we automatically coded as "No". In many other cases though it's unclear whether the task was performed under any kind of supervision.  

```{r}
#| label: fig-task-supervision
#| fig-cap: "Whether sessions were supervised (e.g. in-lab vs. unsupervised online)"
#| fig-width: 8
#| fig-height: 3
make_perc_gg_barplot(clean, "task_supervision")
```

### Other details

Details about participant breaks or management are rare, as are the use of comprehension checks, confidence scoring or pilot testing of the task.  


```{r}
#| label: tbl-admin-details
#| tbl-cap: "Reporting rates for other task administration details"
clean |>
  select(pilot_testing, breaks, comprehension, confidence) |>
  pivot_longer(everything(), names_to = "Detail", values_to = "value") |>
  mutate(
    Detail = case_when(
      Detail == "pilot_testing" ~ "Pilot testing conducted",
      Detail == "breaks" ~ "Breaks offered",
      Detail == "comprehension" ~ "Comprehension checks included",
      Detail == "confidence" ~ "Confidence ratings used"
    ),
    value = ifelse(value %in% c("Not Reported", ""), "Not Reported", "Reported")
  ) |>
  count(Detail, value) |>
  pivot_wider(names_from = value, values_from = n, values_fill = 0) |>
  mutate(perc_reported = round(100 * Reported / (Reported + `Not Reported`), 1)) |>
  rename(`% Reported` = perc_reported) |>
  knitr::kable()
```

:::
