# Transparency {#sec-transparency}

```{r}
#| label: setup
suppressMessages(library(here)) # for pathing to/from project directory
suppressMessages(library(tidyverse)) # for dplyr, ggplot, etc.
suppressMessages(library(janitor)) # for cleaning data
suppressMessages(library(glue)) # for cleaning data

# source scripts
clean <- readRDS(here("data", "clean.rds"))
full <- readRDS(here("data", "full.rds"))

source(here("R", "helpers.R"))
source(here("R", "plots.R"))
```

## Key takeaways

- **Materials are more often shared than data**, but still less than half the time. When items *are* made available, they're most often published directly in the article or its supplementary materials rather than deposited in a dedicated repository.
- **Raw data and analysis code are rarely made available.** The large majority of studies say nothing about whether data or code can be accessed.
- **Preregistration is essentially absent** from this literature — only a handful of studies report having preregistered their design or analysis.
- There's a visible upward trend in materials and data sharing in more recent years, likely reflecting the broader open-science movement, but it starts from a very low baseline.

## Materials availability

```{r}
#| label: fig-item-avail
#| fig-cap: "How (or whether) experimental items were made available"
#| fig-width: 8
#| fig-height: 4
make_perc_gg_barplot(clean, "item_avail")
```

## Data and code availability

```{r}
#| label: fig-data-avail
#| fig-cap: "How (or whether) raw data and/or analysis code were made available"
#| fig-width: 8
#| fig-height: 4
make_perc_gg_barplot(clean, "data_avail")
```


## Preregistration

```{r}
#| label: tbl-prereg
#| tbl-cap: "Preregistration is almost never reported in this literature"
clean |>
  count(preregistration) |>
  mutate(perc = round(100 * n / sum(n), 1),
         text = glue("{n} ({perc}%)")) |>
  select(preregistration, text) |>
  pivot_wider(names_from = preregistration, values_from = text) |>
  knitr::kable()
```
